Summarize twas

### 2020-08-04 analysis/index.Rmd
change sa2 grid

>>>>

## Summarize cTWAS results
We compare cTWAS results with other methods, summarizing results over multiple simulation runs.

  * [Using UKbiobank data (chr22), fusion weights Adipose](simulation-multi-ukbchr22-gtex.adipose.html)
  * [Using UKbiobank data (chr22), fusion weights Adipose, setting 2](simulation-multi-ukbchr22-gtex.adipose2.html)
  * [Using UKbiobank data (chr17:22), fusion weights Adipose](simulation-multi-ukbchr17to22-gtex.adipose.html) 
  * [Using UKbiobank data (chr17:22), fusion weights Adipose, using smaller sa2 grid in mr.ash](simulation-multi-ukbchr17to22-gtex.adipose_sa2_v2.html)

### 2020-08-04 analysis/simulation-multi-ukbchr22-gtex.adipose2.Rmd
change sa2 grid

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
tags <- paste0('20200616-8-', 1:9)
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
```

@@

  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
}
```
# Mr.ash2 parameter estimation
Results for 9 simulations runs, using different initiate and update strategy
```{r param function}
show_param <- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  parf <- lapply(f, '[[', "par")
  param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
  knitr::kable(param)
}
```

@@

```{r param 1}
show_param(tag2s[1])
```

@@

```{r param 2}
show_param(tag2s[2])
```

@@

```{r param 3}
show_param(tag2s[3])
```

@@

```{r param 4}
show_param(tag2s[4])
```

@@

```{r rpip 1, fig.height=4, fig.width=10}
f <- get_files(tag= tags[1], tag2 = tag2s[1])
a <- read.table(f[["rpip"]], header = T)
plot(a$p0, a$rPIP, pch =19, col ='salmon', xlab = "position", ylab= "Sum of PIP")
grid()
```

# PIP scatter plot
mr.ash2s PIP vs. susie PIP.
```{r scatterplot}
scatter_plot_PIP<- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

@@

  fig <- subplot(fig1, fig2, titleX = TRUE, titleY = T, margin = 0.1)
  fig
}
```
## NULL; expr-snp; expr-snp 
```{r sca 1, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[1])
```

@@

```{r sca 2, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[2])
```

@@

```{r sca 3, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[3])
```

@@

```{r sca 4, fig.height=4, fig.width=10}
scatter_plot_PIP(tag2s[4])
```

# ROC curve
```{r roc}
ROC_plot<- function(tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

@@

## NULL; expr-snp; expr-snp 
```{r roc 1, fig.height=4, fig.width=4}
ROC_plot(tag2s[1])
```

@@

```{r roc 2, fig.height=4, fig.width=4}
ROC_plot(tag2s[2])
```

@@

```{r roc 3, fig.height=4, fig.width=4}
ROC_plot(tag2s[3])
```

@@

```{r roc 4, fig.height=4, fig.width=4}
ROC_plot(tag2s[4])
```

>>>>

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
tags <- paste0('20200616-8-', 1:9)
tagglob <- '20200616-8-*'
tagextr <- '20200616-8-\\d+'
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
```

@@

  return(tibble::lst(par, rpip, gmrash, ggwas, smrash, sgwas, gsusie, ssusie))
}

get_tags <- function(globpattern, extrpattern, tag2){
  lapply(lapply(get_files(globpattern, tag2), Sys.glob), str_extract, pattern = extrpattern)
}
```
# Mr.ash2 parameter estimation
Results for 9 simulations runs, using different initiate and update strategy
```{r param function}
show_param <- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  parf <- lapply(f, '[[', "par")
  param <- do.call(rbind, lapply(parf, function(x) t(read.table(x))[2:1,]))
  truth <- param[1:(nrow(param)/2)*2-1,]
  est <- param[1:(nrow(param)/2)*2,]
  outdt <- matrix(0, ncol = 2*ncol(param), nrow = nrow(param)/2)
  outdt[,c(1,3,5,7)] <- truth
  outdt[,c(2,4,6,8)] <- est
  outdt <-cbind(1:nrow(outdt),outdt)
  colnames(outdt) <- c("Simulation#", paste0(rep(c("Truth","Est."),4)))
  knitr::kable(outdt) %>%
  kable_styling("striped") %>%
  add_header_above(c(" " = 1, "Gene.pi1" = 2, "Gene.PVE" = 2, "SNP.pi1" = 2, "SNP.PVE" =2))
}
```

@@

```{r param 1}
show_param(tags = tags, tag2 = tag2s[1])
```

@@

```{r param 2}
show_param(tags = tags, tag2 = tag2s[2])
```

@@

```{r param 3}
show_param(tags = tags, tag2 = tag2s[3])
```

@@

```{r param 4}
show_param(tags = tags, tag2 = tag2s[4])
```

@@

```{r rpip 1, fig.height=4, fig.width=10}
f <- get_files(tag= tags[1], tag2 = tag2s[1])
a <- read.table(f[["rpip"]], header = T)
par(mar=c(5, 4, 4, 6) + 0.1)
with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
par(new = T)
with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
axis(side = 4)
mtext(side = 4, line = 3, 'No. causal signals')
legend("topleft",
       legend=c("Mr.ASH PIP", "# Causal"),
       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
grid()
```

# PIP calibration
We run 100 simulations and combine results.
```{r caliplot}
#' s is pip or fdr.
cp_plot <- function(s, ifcausal, mode = c("PIP", "FDR"), main = mode[1]){
  # ifcausal:0,1
  a_bin <- cut(s, breaks= seq(0,1, by=0.1))
  if (mode == "PIP") {
     expected = c(by(s, a_bin, FUN = mean))
     observed = c(by(ifcausal, a_bin, FUN = mean))
  } else if (mode == "FDR"){
     expected = c(by(s, a_bin, FUN = mean))
     observed = 1 - c(by(ifcausal, a_bin, FUN = mean))
  }
  plot(expected, observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main)
  lines(x = c(0,1), y = c(0,1), col ="red")
}

caliPIP_plot <- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  res <- res[complete.cases(res),]
  res <- rename(res, c("PIP" = "mr.ash_PIP", "pip" = "SUSIE_PIP", "pip.null" = "SUSIE_PIP_null", "pip.w0" = "SUSIE_PIP_w0"))
  par(mfrow=c(1,4), mar=c(5, 6, 4, 1))
  cp_plot(res$mr.ash_PIP, res$ifcausal, main = "Mr.ash PIP")
  cp_plot(res$SUSIE_PIP, res$ifcausal, main = "SUSIE PIP")
  cp_plot(res$SUSIE_PIP_null, res$ifcausal, main = "SUSIE PIP null")
  cp_plot(res$SUSIE_PIP_w0, res$ifcausal, main = "SUSIE PIP weighted null")
}

caliFDR_plot <- function(tags, tag2){

  f <- lapply(tags, get_files, tag2 = tag2)
  gwasf <- lapply(f, '[[', "ggwas")
  names(gwasf) <- tags

  .tagname <- function(x, flist, colnames = NULL){
    a <- read.table(flist[[x]], header =T)
    if (!is.null(colnames)){
      colnames(a) <- colnames
    }
    a[, "name"] <- paste0(x, ":", a[, "name"])
    a
  }

  .addFDR <- function(res){
    res$FDR <- p.adjust(res$PVALUE, method = "fdr")
    res
  }

  gwasres <- do.call(rbind, lapply(lapply(tags, .tagname, flist = gwasf, 
                                   colnames =  c("chr",	"p0",	"p1", "name", 
                                                 "Estimate", "Std.Error", "t-value", "PVALUE")), .addFDR))

  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

  res <- merge(res, gwasres, by = "name", all = T)

  res <- res[complete.cases(res),]

  cp_plot(res$FDR, res$ifcausal, mode ="FDR", main = "TWAS FDR")
  cat("FDR at bonferroni corrected p = 0.05: ", 1 - mean(res[gwasres$PVALUE < 0.05 /dim(res)[1], "ifcausal"]))
}
```

## NULL; expr-snp; expr-snp 
```{r cali 1, fig.height=4, fig.width=14}
tag2 = "zeroes-es"
tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
res <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
```
```{r cali 1fdr, fig.height=4, fig.width=4}
caliFDR_plot(tags = tags_ext, tag2 = tag2)
```

## Lasso; expr-snp; expr-snp 
```{r cali 2, fig.height=4, fig.width=14}
caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
```

```{r cali 2fdr, fig.height=4, fig.width=4}
caliFDR_plot(tags = tags_ext, tag2 = "lassoes-es")
```

# PIP scatter plot
mr.ash2s PIP vs. susie PIP.
```{r scatterplot}
scatter_plot_PIP<- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

@@

  fig <- subplot(fig1, fig2, titleX = TRUE, titleY = T, margin = 0.1)
  fig
}
```
## NULL; expr-snp; expr-snp 
```{r sca 1, fig.height=4, fig.width=10}
scatter_plot_PIP(tags = tags, tag2 = tag2s[1])
```

@@

```{r sca 2, fig.height=4, fig.width=10}
scatter_plot_PIP(tags = tags, tag2 = tag2s[2])
```

@@

```{r sca 3, fig.height=4, fig.width=10}
scatter_plot_PIP(tags = tags, tag2 = tag2s[3])
```

@@

```{r sca 4, fig.height=4, fig.width=10}
scatter_plot_PIP(tags = tags, tag2 = tag2s[4])
```

# ROC curve
```{r roc}
ROC_plot<- function(tags, tag2){
  f <- lapply(tags, get_files, tag2 = tag2)
  mrashf <- lapply(f, '[[', "gmrash")
  names(mrashf) <- tags

@@

## NULL; expr-snp; expr-snp 
```{r roc 1, fig.height=4, fig.width=4}
ROC_plot(tags = tags, tag2 = tag2s[1])
```

@@

```{r roc 2, fig.height=4, fig.width=4}
ROC_plot(tags = tags, tag2 = tag2s[2])
```

@@

```{r roc 3, fig.height=4, fig.width=4}
ROC_plot(tags = tags, tag2 = tag2s[3])
```

@@

```{r roc 4, fig.height=4, fig.width=4}
ROC_plot(tags = tags, tag2 = tag2s[4])
```
